\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{comment}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{enumitem}
\usepackage{scrextend}

\newcommand{\iimplies}{\mbox{ IMPLIES }}
\newcommand{\oor}{\mbox{ OR }}
\newcommand{\aand}{\mbox{ AND }}
\newcommand{\nnot}{\mbox{ NOT }}
\newcommand{\iiff}{\mbox{ IFF }}
\newcommand{\xxor}{\mbox{ XOR }}
%\newcommand{\algorithmicbreak}{\textbf{break}}
%\newcommand{\Break}{\State \algorithmicbreak}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\begin{center}
{\bf \Large \bf CSC265H Assignment 6}
\end{center}

\noindent
Yibin Zhao\\
1002996261\\
Dicussed with Jiayi Zhang\\
NO EXTRA MATERIAL CONSULTED\\

\begin{comment}
The median of a set of numbers is the element with rank $\lfloor (n+1)/2 \rfloor$.
\end{comment}

\begin{enumerate}
	
	\begin{comment}
	Prove that any decesion tree using 2-way comparisons that finds the median of $n$ distinct numbers 
	also determines the set of numbers greater than the median and the set of numbers smaller than the median.
	\end{comment}

	\item

	I will prove this propesition by proving the following stronger theorem:
	
	\begin{theorem}
	Any decision tree using 2-way comparison that finds that $k$th largest element in $n$ distinct numbers
	also determines the set of numbers greater than the $k$th largest element and
	the set of numbers smaller than the $k$th largest element, where $k$ is any integer from 1 to $n$.
	\end{theorem}

	\begin{lemma}
	For a decision tree, the result of a 2-way = comparison is consistent for any input of size $n$.
	\end{lemma}

	\begin{proof}
	Consider a comparison $A[i] = A[j]$ in the decision tree, where $i, j \in \{1, 2, \cdots, n\}$.

	If $i = j$, then the comparison always return T. 
	Otherwise, it always returns F regardless of the input, since all numbers are distinct.
	\end{proof}

	Thus, the result of a $=$ comparison does not impliy anything about the structure of the input.

	Therefore, we can safely assume all comparisons below are 2-way $<, >, \leq, \geq$ comparisons, 
	unless I explicitly state that it is a $=$ comparison.

	\begin{lemma}
	If $k > 1$, then there is at least one comparison between the $k$th largest number and the $k-1$ largest number.
	\end{lemma}
	
	\begin{proof}
	Let A be any array with distinct numbers of size $n$.
	Since $k>1$, $n\geqk>1$.
	Let $p, q \in \{1, 2, \ldots, n\}$, such that $A[p]$ has rank $k-1$ and $A[q]$ has rank $k$.

	Consider an arbitrary decision tree with 2-way comparison that gives the index of the $k$th largest
	number in this array.
	Assume that there is no comparison between $A[p]$ and $A[q]$ in the path for this input $A$.
	Then, let $B[i] = A[i]$ for all $i$ not equal to $p$ or $q$, and let $B[p] = A[q]$ and $B[q] = A[p]$.

	Consider a comparison between $B[i]$ and $B[j]$, where $i \neq j$.
	By the assumption above, $\{i, j\} \neq \{p, q\}$. \\
	If neither $i, j$ equal to $p$ or $q$, the result is equal to the same comparison between $A[i]$ and $A[j]$. \\
	If one of $i, j$ is equal to $p$ or $q$. 
	Without loss of generality, suppose $j$ is not equal to either $p$ or $q$. 
	Then, if $A[j]$ has a rank smaller than $k-1$, then $A[j] = B[j] < A[p] = B[q]$ and $B[j] < A[q] = B[p]$. 
	This implies that the comparison between $B[i]$ and $B[j]$ is consistent with the comparison between $A[i]$ and $A[j]$.
	Otherwise, $A[j]$ has a rank greater than $k$, then $A[j] = B[j] > A[p] = B[q]$ and $B[j] > A[p] = B[q]$. 
	Thus, its also consistent in this case.
	Similarly, we can obtain that the result of this comparison is also consistent for $i$ not equal to either $p$ or $q$. \\
	
	Thus, the output is consistent for $A$ and $B$.
	However, the output is not correct for $B$. 
	This is a contradiction.

	Therefore, there is at least one comparison between $A[p]$ and $A[q]$.
	By generalization, the lemma is true.
	\end{proof}

	\begin{corrolary}
	If $k < n$, then there is at least one comparison between the $k$th largest number and the $k+1$ largest number.
	\end{corrolary}

	\begin{proof}
	Use similar approach as the lemma, not stating here.
	\end{proof}

	We can now prove the theorem.
	\begin{proof}
	Let $P(n)$ be the predicate that "the theorem is true for size $n$ and all $k$".
	Want to prove that $\forall n \in \mathbb{Z}^+. P(n)$. 

	\textbf{Base Case:} $n=1$
	\begin{addmargin}[1em]{0em}
		Then $k=1$. 
		The decision tree always return the one and only element in this array.
		The set smaller than that and the set greater than that element is always empty.
		Thus, $P(1)$.
	\end{addmargin}

	\textbf{Inductive Step:}
	\begin{addmargin}[1em]{0em}
		Suppose that $P(n)$ is true for some $n \in \mathbb{Z}^+$. 
		Want to prove that $P(n+1)$ is true.

		If $k = 1$, then it is trivial,
		since all other numbers except the output are in the set of numbers greater than the output,
		and the set of numbers smaller than that is alwasy empty.

		If $k = n+1$, then it is also trivial, 
		since all other numbers except the output are in the set of numbers smaller than the output, 
		and the set of numbers greater than that is always empty.

		Now consider $1 < k < n+1$, if that case exists.
		Suppose there is a decision tree $T$ that solve the problem for $n+1$ and $k$.

		Consider an arbitrary array $A$ of size $n+1$ that satisfy $\forall i \in \{1, \cdots, n\}. A[n+1] > A[i]$.
		Then the $k$th largest element in $A[1 \ldots n+1]$ is also 
		the $k$th largest element in $A[1 \ldots n]$.

		Define that following procedure on tree $T$:
		\begin{addmargin}[1em]{0em}
			If there exists a node that contants a comparison between $A[n+1]$ and $A[i]$, 
			where $i \in \{1, \cdots, n\}$.

			If this comparison returns T, then replace this node by its T child and delete the F subtree. \\
			Otherwise, replace this node by its F child and delete the T subtree.

			Repeat above steps until there is no such node in this tree.
		\end{addmargin}

		Let $T'$ be the new tree after this procedure.
		Notice that $T'$ does not depend on the actual input array $A$.
		It only depends on $T$ and the assumption of $A$, on which $A[n+1]$ is the largest element.

		Thus, the result of $T'$ on input $A[1 \ldots n]$ is consistent with the result of $T$ on input $A[1 \ldots n+1]$.
		Since for any arbitrary $B[1 \ldots n]$, we can always append a number that is greater than 
		all elements in $B$, we can claim that $A[1 \cdots n]$ is an arbitrary array on size $n$. 
		Therefore, $T'$ is a decision tree that solves the problem with $n$ and $k$.

		By the inductive assumption, $T'$ can determine the set of numbers smaller than the $k$th largest number
		and the set of numbers greater than the $k$th largest number for problem of size $n$.
		



	\end{addmargin}
	
	\end{proof}


	

	\begin{comment}
	Prove and $\Omega(n)$ lower bound on the worst case complexity of finding the median of $n$ numbers 
	using the decision tree with 2-way comparisons.
	\end{comment}

	\item By 1, we can make the output of the decision tree be the index of the median and the set of indecies of elements greater than the median and the set of indecies for smaller ones. \\
	Then, the total number of possible output for an array of size $n$ is
	\begin{align*}
		& n \cdot 
		\begin{pmatrix}
			n-1 \\
			\lfloor (n+1)/2 \rfloor - 1
		\end{pmatrix} \\
		=& \frac{n^{(\lfloor (n+1)/2 \rfloor)}}{\lfloor (n-1)/2 \rfloor!} \\
		\geq& \frac{n^{(\lfloor n/2 \rfloor)}}{\lfloor n/2 \rfloor!} \\
		=& \prod_{i=1}^{\lfloor n/2 \rfloor}\frac{n-i+1}{\lfloor n/2 \rfloor -i+1} \\
		\geq& 2^{\lfloor n/2 \rfloor}
	\end{align*}
	Since the information theory lower bound $\log_2(2^{\lfloor n/2 \rfloor}) = \left\lfloor \frac{n}{2} \right\rfloor \in \Omega(n)$, 
	$\Omega(n)$ is the lower bound of the worst case complexity.

\end{enumerate}

\end{document}
